---
title: "NBA Point Guards"
output: html_document
---

```{r, child="children/SETTINGS-knitr.txt"}
```

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(tibble)
```

## Read data

```{r}
minutes = 1500

data <- read_csv("basketball-reference.csv")
data <- data %>% replace_na(list("3P%" = 0))

pointguards <- data %>% filter(Pos == "PG", MP > minutes) %>% column_to_rownames("Player") %>% select(-Rk, -Pos, -Age, -Tm, -G, -GS, -MP, -X30, -FGA, -"3PA", -"2PA", -FTA, -TRB)
dim(pointguards)
shootingguards <- data %>% filter(Pos == "SG", MP > minutes) %>% column_to_rownames("Player") %>% select(-Rk, -Pos, -Age, -Tm, -G, -GS, -MP, -X30, -FGA, -"3PA", -"2PA", -FTA, -TRB)
dim(shootingguards)
smallforwards <- data %>% filter(Pos == "SF", MP > minutes) %>% column_to_rownames("Player") %>% select(-Rk, -Pos, -Age, -Tm, -G, -GS, -MP, -X30, -FGA, -"3PA", -"2PA", -FTA, -TRB)
dim(smallforwards)
powerforwards <- data %>% filter(Pos == "PF", MP > minutes) %>% column_to_rownames("Player") %>% select(-Rk, -Pos, -Age, -Tm, -G, -GS, -MP, -X30, -FGA, -"3PA", -"2PA", -FTA, -TRB)
dim(powerforwards)
centers <- data %>% filter(Pos == "C", MP > minutes) %>% column_to_rownames("Player") %>% select(-Rk, -Pos, -Age, -Tm, -G, -GS, -MP, -X30, -FGA, -"3PA", -"2PA", -FTA, -TRB)
dim(centers)
```

## Centers

### Compute the archetypes

```{r}
library(biaa)

set.seed(2020)

scaled_centers = scale(as.matrix(centers))

k = 2:6
c = 2:6

arch = biaa(scaled_centers, k, c, nrep = 30, maxiter = 200000)
```

#### Plotting the RSS
```{r}
rssplot(arch)
```

#### Getting the best model

```{r}
bk = 3
bc = 3
barch = getmodel(arch, k = bk, c = bc)
```


### Results

```{r}
barch$rss

barch$archetypes
```

```{r}
ogammas = as_tibble(t(barch$gammas)) %>% add_column(VARIABLE = colnames(centers))

thresh = 0.5
for (i in 1:bc) {
  print(ogammas %>% filter(.[[i]] > thresh) %>% select(VARIABLE, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))

}
```


```{r}
oalphas = as_tibble(barch$alphas) %>% add_column(PLAYER = row.names(centers))

thresh = 0.6
for (i in 1:bk) {
  print(oalphas %>% filter(.[[i]] > thresh) %>% select(PLAYER, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))
}
```


## Power Forwards

### Compute the archetypes

```{r}
library(biaa)

set.seed(2020)

scaled_powerforwards = scale(as.matrix(powerforwards))

k = 2:6
c = 2:6

arch = biaa(scaled_powerforwards, k, c, nrep = 30, maxiter = 200000)
```

#### Plotting the RSS
```{r}
rssplot(arch)
```


#### Getting the best model

```{r}
bk = 4
bc = 4
barch = getmodel(arch, k = bk, c = bc)
```


### Results

```{r}
barch$rss

barch$archetypes
```

```{r}
ogammas = as_tibble(t(barch$gammas)) %>% add_column(VARIABLE = colnames(powerforwards))

thresh = 0.5
for (i in 1:bc) {
  print(ogammas %>% filter(.[[i]] > thresh) %>% select(VARIABLE, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))

}
```


```{r}
oalphas = as_tibble(barch$alphas) %>% add_column(PLAYER = row.names(powerforwards))

thresh = 0.6
for (i in 1:bk) {
  print(oalphas %>% filter(.[[i]] > thresh) %>% select(PLAYER, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))
}
```

## Small Forwards

### Compute the archetypes

```{r}
library(biaa)

set.seed(2020)

scaled_smallforwards = scale(as.matrix(smallforwards))

k = 2:6
c = 2:6

arch = biaa(scaled_smallforwards, k, c, nrep = 30, maxiter = 200000)
```

#### Plotting the RSS
```{r}
rssplot(arch)
```

#### Getting the best model

```{r}
bk = 4
bc = 4
barch = getmodel(arch, k = bk, c = bc)
```


### Results

```{r}
barch$rss

barch$archetypes
```

```{r}
ogammas = as_tibble(t(barch$gammas)) %>% add_column(VARIABLE = colnames(smallforwards))

thresh = 0.5
for (i in 1:bc) {
  print(ogammas %>% filter(.[[i]] > thresh) %>% select(VARIABLE, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))

}
```


```{r}
oalphas = as_tibble(barch$alphas) %>% add_column(PLAYER = row.names(smallforwards))

thresh = 0.6
for (i in 1:bk) {
  print(oalphas %>% filter(.[[i]] > thresh) %>% select(PLAYER, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))
}
```

## Shooting Guards

### Compute the archetypes

```{r}
library(biaa)

set.seed(2020)

scaled_shootingguards = scale(as.matrix(shootingguards))

k = 2:6
c = 2:6

arch = biaa(scaled_shootingguards, k, c, nrep = 30, maxiter = 200000)
```

#### Plotting the RSS
```{r}
rssplot(arch)
```

#### Getting the best model

```{r}
bk = 3
bc = 3
barch = getmodel(arch, k = bk, c = bc)
```


### Results

```{r}
barch$rss

barch$archetypes
```

```{r}
ogammas = as_tibble(t(barch$gammas)) %>% add_column(VARIABLE = colnames(shootingguards))

thresh = 0.5
for (i in 1:bc) {
  print(ogammas %>% filter(.[[i]] > thresh) %>% select(VARIABLE, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))

}
```


```{r}
oalphas = as_tibble(barch$alphas) %>% add_column(PLAYER = row.names(shootingguards))

thresh = 0.6
for (i in 1:bk) {
  print(oalphas %>% filter(.[[i]] > thresh) %>% select(PLAYER, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))
}
```

## Point Guards

### Compute the archetypes

```{r}
library(biaa)

set.seed(2020)

scaled_pointguards = scale(as.matrix(pointguards))

k = 2:6
c = 2:6

arch = biaa(scaled_pointguards, k, c, nrep = 30, maxiter = 200000)
```

#### Plotting the RSS
```{r}
rssplot(arch)
```

#### Getting the best model

```{r}
bk = 5
bc = 4
barch = getmodel(arch, k = bk, c = bc)
```


### Results

```{r}
barch$rss

barch$archetypes
```

```{r}
ogammas = as_tibble(t(barch$gammas)) %>% add_column(VARIABLE = colnames(pointguards))

thresh = 0.5
for (i in 1:bc) {
  print(ogammas %>% filter(.[[i]] > thresh) %>% select(VARIABLE, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))

}
```


```{r}
oalphas = as_tibble(barch$alphas) %>% add_column(PLAYER = row.names(pointguards))

thresh = 0.5
for (i in 1:bk) {
  print(oalphas %>% filter(.[[i]] > thresh) %>% select(PLAYER, SIMILARITY = i) %>% arrange(desc(SIMILARITY)))
}
```